---
title: Book review statistics
author: Samal Abdikerimova, Ravi Donepudi, Carlos Salinas, and Xiran Yu
output:
  html_document
---

# Book review statistics

```{r setup, include = FALSE}
library("tidyr")
library("broom")
library("tidyverse")
library("rgoodreads")
library("forcats")
```

[rgoodreads]: https://github.com/Famguy/rgoodreads
[goodreads-kaggle]: https://www.kaggle.com/jealousleopard/goodreadsbooks/
## Ideas

1. Look at relationship between reviews and pages.
2. Filter by language.
3. Relationship between popularity and gender.
4. Language and popularity.
5. Genre popularity through the years.
6. Publisher relationship with rating.
7. Publisher and gender relationship.

Note that there is an Goodreads API and that somebody has written an R interface
to it, however we do not have enough time to explore it and get it to do as we
want. What we will do instead is use the data set found on Kaggle [here][goodreads-kaggle]together with the Goodreads API to answer some of the questions above.

First we import the data and take a good look.

```{r import-books_csv, message = FALSE, warning = FALSE}
books <- read_csv("data/books.csv") %>%
    rename(rating = average_rating,
           total_ratings = ratings_count,
           lang = language_code,
           pages = num_pages)
books
```

Let's get more of an idea of what we are working with. Since we speak English we'd like to focus solely on English language books. But first let's see how many of them we have

```{r, message = FALSE, warning = FALSE}
books %>%
    subset(is.na(as.numeric(lang))) %>%
    subset(!is.na(lang)) %>%
    group_by(lang) %>%
    tally() %>%
    arrange(desc(n)) %>%
    slice(1:10) %>%
    ggplot(aes(x = factor(lang, levels = lang), y = n, fill = lang)) +
    geom_bar(stat = "identity") +
    labs(title = "Title number of books by language",
         x = "language",
         y = "number of books") +
    theme(axis.text.x = element_text(angle=90, hjust=1))
```

Nice, so the overwhelming majority of the data is in `eng`. However, it looks like US and Great Britain version of books fall under a different language tag (and if we search hard Canadian too). For our purposes `"en-CA", `"en-GB"`, and `en-US` are the same so let us clean up the data a little 

```{r eng_replace, message = FALSE, warning = FALSE, eval = FALSE}
books$lang <- books %>%
    pull(lang) %>%
    str_replace_all("en*(g|-GB|-US|-CA)", "eng")
```

Another way to deal with this problem is simply to ignore these variants by filtering them out of the data set. Now, since the overwhelming majority of the books in this data set are in English, we will focus our attention on `"eng"`.

```{r english books, message = FALSE, warning = FALSE}
eng_books <- books %>% filter(lang == "eng") %>% filter(rating >= 3)
```

One thing we might one to look at first is the distribution of ratings for English books. We do this as follows 

```{r plot ratings, message = FALSE, warning = FALSE}
eng_books %>%
    group_by(rating) %>%
    summarize(total_books = n()) %>%
    ungroup() %>%
    ggplot(aes(x = rating)) +
    geom_bar(aes(y = total_books/1.000001, fill = rating),
             stat = "identity") +
    labs(title = "Total number of books per rating",
         x = "average rating",
         y = "number of books")
```

OK, to make better sense of this data let's take a look at how many people are reading these books

and who are the most prolific writers

```{r prolific writers, message = FALSE, warning = FALSE}
prolific_writer_names <- eng_books %>%
    group_by(authors) %>%
    summarize(count = n()) %>%
    arrange(desc(count)) %>%
    ungroup() %>%
    slice(1:10) %>%
    pull(authors)

prolific_writers <- eng_books %>% filter(authors %in% prolific_writer_names)
```

Stephen King—that was sort of expected—and who are the top rated authors

```{r naive top authors, message = FALSE, warning = FALSE}
eng_books %>%
    group_by(authors) %>%
    summarize(average_rating = sum(rating)/n()) %>%
    arrange(desc(average_rating)) %>%
    ungroup() %>%
    slice(1:10)
```

who are these people? So highly rated authors are not necessarily notable ones, perhaps there is a better way to measure top authors.

```{r good books, message = FALSE, warning = FALSE}
eng_books %>%
    (function(...) {
        tmp <- tibble(...)
        tmp$authors <- tmp$authors %>%
            str_replace_all("Margaret Weis/Tracy Hickman",
                            "Weis/Hickman")
        tmp}) %>%
    group_by(authors) %>%
    summarize(book_count = n(),
              sum_rating = sum(rating)) %>%
    arrange(desc(sum_rating)) %>%
    slice(1:10) %>%
    ggplot(aes(factor(authors, levels = authors), sum_rating, fill = authors)) +
    geom_col() +
    coord_flip() +
    labs(title = "Author rankings",
         y = "author",
         x = "sum rating") +
    scale_fill_brewer(palette="Spectral")
```

```{r average rating, message = FALSE, rating = FALSE}
prolific_writers %>%
    group_by(authors) %>%
    summarize(avg_rat = sum(rating)/n()) %>%
    ungroup() %>%
    ggplot(aes(x = authors, y = avg_rat, fill = authors)) +
    geom_bar(stat = "identity")
```

```{r, message = FALSE, warning = FALSE}
eng_books <- eng_books %>%
    mutate(publication_year = as.numeric(str_sub(publication_date, -4)))
```

```{r publications per year, message = FALSE, warning = FALSE}
eng_books %>% group_by(publication_year) %>%
    summarize(total_books = n()) %>%
    ungroup() %>%
    ggplot(aes(x = publication_year, y = total_books)) +
    geom_point() +
    xlim(1900, 2020) +
    labs(x = "publication year",
         y = "books published")
```

```{r, message = FALSE, rating = FALSE}
books_per_year <- eng_books %>% group_by(publication_year) %>%
    summarize(total_books = n()) %>%
    arrange(desc(total_books)) %>%
    ungroup()
```

```{r, warning = FALSE, rating = FALSE}
model_publish_books <- books_per_year %>%
    filter(publication_year <= 2006) %>%
    lm(log(total_books) ~ publication_year, .)

filtered_model <- books_per_year %>%
    filter(publication_year > 2006) %>%
    lm(log(total_books) ~ publication_year, .)

model_publish_books %>% tidy()
filtered_model %>% tidy()

estimate_2020 <-
    predict(model_publish_books,
            newdata = data.frame(publication_year = seq(2015,2020)))
```

```{r, message = FALSE, rating = FALSE}
stephen_king <- books %>% filter(authors == "Stephen King" &
                                 title != "Blood and Smoke" &
                                 title != "LT's Theory of Pets")
sk_model <- lm(rating ~ pages, stephen_king)
sk_model %>% tidy()

predict(sk_model,
        newdata = data.frame(pages = c(300)))
```

```{r, message = FALSE, rating = FALSE}
stephen_king %>% summarize(mean = mean(pages))
```

```{r, message = FALSE, warning = FALSE}
eng_books$publisher <- factor(eng_books$publisher)
```

```{r, message = FALSE, warning = FALSE}
eng_books %>% pull(publisher) %>% levels()
```

```{r, message = FALSE, warning = FALSE }
top_10_publishers <- eng_books %>%
    group_by(publisher) %>%
    summarize(total_books = n()) %>%
    arrange(desc(total_books)) %>%
    slice(1:10) %>%
    pull(publisher)
```

```{r, message = FALSE, warning = FALSE}
eng_books_tp <- eng_books %>%
    filter(publisher %in% top_10_publishers)
```
